---
title: "Message ET Code"
author: "Niklas D."
format: html
editor: visual
---

# Import packages

```{r setup}
library(tidyverse)

library(cmdstanr)
library(crosstable)
library(easystats)
library(glue)
library(readxl)
library(vtree)

ukrain_start <- as.Date("2022-02-24")
hamas_start <- as.Date("2023-10-07")

colors_pre_post <- c("#97BF0D", "#E6444F")
```

# Import data

```{r}
message_complete_df <- read_xlsx("data/Datensatz_aktuell.xlsx") |>
  mutate(across(everything(), ~ replace_na(., 0))) #Fill NAs

message_complete_df_long <- message_complete_df |> 
  rename("A17" = "A17 miscell") |> 
  pivot_longer(c(B2, B11, B12, B13, A11, A12, A13, A14, A15, A16, A17, A21, A22), names_to = "name", values_to = "present") |> 
  separate_wider_position(name, widths = c(motive = 1, target = 1, topic = 1), too_few = "align_start") |> 
  mutate(target = as.numeric(target),
         topic = as.numeric(topic)) |> 
  
  #Tranlate motive
  mutate(motive = case_when(motive == "A" ~ "Information",
                            motive == "B" ~ "Interaction")) |> 
  
  #Translate target
  #Means different things depending on motive
  
  mutate(target = case_when(motive == "Information" & target == 1 ~ "ET",
                             motive == "Information" & target == 2 ~ "Humans",
                             
                             motive == "Interaction" & target == 1 ~ "Affiliation",
                             motive == "Interaction" & target == 2 ~ "Distance")) |> 
  #Translate topic
  #Means different things depending on motive

 mutate(topic = case_when(motive == "Information" & target == "ET" & topic == 1 ~ "Existence",
                           motive == "Information" & target == "ET" & topic == 2 ~ "Localization",
                           motive == "Information" & target == "ET" & topic == 3 ~ "Environment",
                           motive == "Information" & target == "ET" & topic == 4 ~ "Physiology",
                           motive == "Information" & target == "ET" & topic == 5 ~ "Psychology",
                           motive == "Information" & target == "ET" & topic == 6 ~ "Knowledge",
                           motive == "Information" & target == "ET" & topic == 7 ~ "Misc",
                          
                           motive == "Information" & target == "Humans" & topic == 1 ~ "Individual",
                           motive == "Information" & target == "Humans" & topic == 2 ~ "Collective",
                           
                           motive == "Interaction" & target == "Affiliation" & topic == 1 ~ "Greetings",
                           motive == "Interaction" & target == "Affiliation" & topic == 2 ~ "Contact",
                           motive == "Interaction" & target == "Affiliation" & topic == 3 ~ "Help",
                          
                           motive == "Interaction" & target == "Distance" & is.na(topic) ~ NA)) |> 
  filter(present == 1) |> 
  mutate(pre_ukr = case_when(Datum < as.Date("2022-02-24") ~ "before",
                             Datum >= as.Date("2022-02-24") ~ "after"),
         pre_ukr = factor(pre_ukr, levels = c("before", "after")),
         pre_hamas = case_when(Datum < as.Date("2023-10-07") ~ "before",
                             Datum >= as.Date("2023-10-07") ~ "after"),
         pre_hamas = factor(pre_hamas, levels = c("before", "after")))
  

#Check uncategorized

message_complete_df |> 
  mutate(sumscore = rowSums(across(B2:`A17 miscell`), na.rm = TRUE)) |> 
  filter(sumscore == 0)
```

# Descriptive table about categories

Messages were coded in a way that allowed mutliple categories to apply to each message. 606 Messages, yielded 806 utterances. Of these 806 utterances ....

```{r}
tree_interaction <- vtree(message_complete_df_long |> filter(motive == "Interaction"),
      "motive target topic", folder = "data")

tree_information <- vtree(message_complete_df_long |> filter(motive == "Information"),
      "motive target topic")

vtree::grVizToPNG(tree_interaction, folder = "figures")
vtree::grVizToPNG(tree_information, folder = "figures")
```

# Sasha

Question:

Does the occurrence of specific message topics change after the 24.2.2022 (Russia invades ukraine - pre_ukr) and after 7.10.2023 (Hamas attacks Israel - pre_hamas).

Special focus on the War/ecocrisis theme, but also the affiliation topic. War/ecocrisis are overarching categories and not within the hierarchical structure as seen in the "figures" folder.

Dates with no data points do not necessarily have 0 messages, but can also have 0 non-transcribed messages.

It is unclear which time frame we should use (7, 14, 21 days prior/past?) Date of occurrence is "past"


```{r}
tagged_messages_df <-
  message_complete_df_long |>
  mutate(Tag = glue::glue("{motive}-{target}-{topic}")) |>
  mutate(Tag = factor(Tag),
         Tag = fct_recode(Tag, "Interaction-Distance" = "Interaction-Distance-NA")) |>
  group_by(Datum, Tag) |>
  summarise(N = n(), .groups = "drop") |>
  complete(Datum, Tag, fill = list("N" = 0)) |>
  group_by(Datum) |>
  mutate(Ntotal = sum(N),
         P = N / Ntotal) |>
  ungroup()

war_messages_df <-
  message_complete_df_long |>
    group_by(Datum) |>
    summarise(N = sum(War),
              Ntotal = n(),
              P = N / Ntotal, 
              .groups = "drop") |>
    add_column(Tag = "War")

eco_messages_df <-
  message_complete_df_long |>
    group_by(Datum) |>
    summarise(N = sum(Ökokrise),
              Ntotal = n(),
              P = N / Ntotal, 
              .groups = "drop") |>
    add_column(Tag = "Ecology")

all_tags <-
  bind_rows(war_messages_df, eco_messages_df, tagged_messages_df) |>
  mutate(Tag = fct_inorder(Tag))

```

Ukrainian war
```{r}
ukrain_tags <-
  all_tags |>
  filter(Datum < as.Date("2023-01-01")) |>
  mutate(Day = as.integer(difftime(Datum, ukrain_start, units = "days")),
         AbsDay = abs(Day)) |>
  mutate(Period  = ifelse(Day < 0, "Pre", "Post"),
         Period = factor(Period, levels = c("Pre", "Post")))

# Looking at Ukrainian war via counts
ggplot(ukrain_tags, aes(x = Datum, y = N, color = Period)) + 
  geom_point(show.legend = FALSE) +
  facet_wrap(. ~ Tag, scales = "free_y")

ggplot(ukrain_tags, aes(x = Datum, y = P, color = Period)) + 
  geom_point(show.legend = FALSE) +
  facet_wrap(. ~ Tag, scales = "free_y")
```

Hamas attack
```{r}
hamas_tags <-
  all_tags |>
  filter(Datum > as.Date("2023-01-01")) |>
  mutate(Day = as.integer(difftime(Datum, hamas_start, units = "days")),
         AbsDay = abs(Day)) |>
  mutate(Period  = ifelse(Day < 0, "Pre", "Post"),
         Period = factor(Period, levels = c("Pre", "Post")))

ggplot(hamas_tags, aes(x = Datum, y = N, color = Period)) + 
  geom_point(show.legend = FALSE) +
  facet_wrap(. ~ Tag, scales = "free_y")

ggplot(hamas_tags, aes(x = Datum, y = P, color = Period)) + 
  geom_point(show.legend = FALSE) +
  facet_wrap(. ~ Tag, scales = "free_y")

```

## Statistical analysis

### Binomial

Binomial model for War/Ecology tags
```{r}
binomial <- cmdstan_model("stan/binomial.stan")

binom_prediction <- function(params, max_days_pre, max_days_post, iteration) {
  bind_rows(
    tibble(Day = seq(-max_days_pre, 0),
           AbsDay = abs(Day),
           P = kilter::inv_logit(params[1] + params[3] * Day),
           Period = "Pre"),
        tibble(Day = seq(0, max_days_post),
           AbsDay = Day,
           P = kilter::inv_logit(params[2] + params[4] * Day),
           Period = "Post")
  ) |>
  mutate(Period = factor(Period, levels = c("Pre", "Post")),
         Iteration = iteration)
}

comparison_for_binomial <- function(tags_df, event, tag){
  crnt_tags_df <-
    tags_df |>
    filter(Tag == tag)
  
  max_days_pre <- abs(min(crnt_tags_df$Day))
  max_days_post <- max(crnt_tags_df$Day)
  max_days <- min(max_days_pre, max_days_post)
  
  
  crnt_stan_data <- list(
    N = nrow(crnt_tags_df),
    Ntag = crnt_tags_df$N,
    Ntotal = crnt_tags_df$Ntotal,
    Day = crnt_tags_df$Day
  )
  
  fit <- binomial$sample(crnt_stan_data, chains = 4, parallel_chains = future::availableCores())
  draws <- fit$draws(format = "df") |> as_tibble()
  
  # comparing predictions for day zero
  day_zero <-
    draws |>
    select(starts_with("alpha")) |>
    rename("Pre" = 1, "Post" = 2) |>
    mutate(across(everything(), ~boot::inv.logit(.)))
  
  day_zero_diff <-
    day_zero |>
    mutate(D = Post - Pre) |>
    summarise(Info = sprintf("Post-Per=%.2f [%.2f, %.2f]\nP(Post>Pre)=%.1f%%", median(D), kilter::lower_ci(D), kilter::upper_ci(D), 100 * mean(D>0)), .groups = "drop") |>
    pull(Info)
  
  day_zero |>
    pivot_longer(everything(), names_to = "Period", values_to = "P") |>
    mutate(Period = factor(Period, levels = c("Pre", "Post"))) |>
  ggplot(aes(x = Period, y = P, color = Period, fill = Period)) +
    geom_boxplot(outlier.shape = NA, show.legend = FALSE, alpha = 0.5) +
    # geom_violin(show.legend = FALSE, alpha = 0.5) +
    labs(title = glue::glue("{event} - {tag}: Day zero predictions"),
         subtitle = day_zero_diff) +
    scale_color_manual(values = colors_pre_post) +
    scale_fill_manual(values = colors_pre_post)
  
  ggsave(glue("figures/{event}-{tag}-DayZero.png"), units = "cm", width = 12, height = 10)
  
  # comparing predictions for N-days pre/post event
  binom_params <- 
    select(draws, starts_with("alpha"), starts_with("beta")) |>
    as.matrix()
  

  
  posterior_predictions <- purrr::map(1:nrow(binom_params), ~binom_prediction(binom_params[., ], max_days_pre, max_days_post, .), .progress = TRUE) |> list_rbind()
  
  posterior_avg <-
    posterior_predictions |>
    group_by(Period, AbsDay) |>
    summarise(Avg = mean(P),
              LowerCI = kilter::lower_ci(P),
              UpperCI = kilter::upper_ci(P),
              .groups = "drop")
  
  
  posterior_pairwise <-
    posterior_predictions |>
    select(-Day) |>
    pivot_wider(names_from = Period, values_from = P) |>
    drop_na() |>
    mutate(D = Post - Pre) |>
    group_by(AbsDay) |>
    summarise(Info = sprintf("%.2f\n%.2f\n%.2f\n%.1f%%", kilter::lower_ci(D), median(D), kilter::upper_ci(D), 100 * mean(D > 0)),
              P = mean(D > 0),
              .groups = "drop") |>
    arrange(AbsDay)
    
    
  
  ggplot(crnt_tags_df |> filter(AbsDay > 0), aes(x = AbsDay, y = P, color = Period))  +
    geom_point(data = posterior_pairwise, color = "black")+
    geom_hline(yintercept = c(0.9, 0.95), linetype = "dashed") +
    geom_ribbon(data = posterior_avg, aes(y = Avg, ymin = LowerCI, ymax = UpperCI, fill = Period), color = NA, alpha = 0.35) +
    geom_line(data = posterior_avg, aes(y = Avg)) +
    geom_point() +
    scale_x_continuous(name = "Day") +
    scale_y_continuous(name = "Probability of message", sec.axis = dup_axis(name = NULL, breaks = c(0.9, 0.95))) +
    scale_color_manual(values = colors_pre_post) +
    scale_fill_manual(values = colors_pre_post) +
    labs(title = glue::glue("{event} - {tag}"))
  
  
  ggsave(glue("figures/{event}-{tag}.png"), units = "cm", width = 24, height = 14)
}

events_tags <- list("Ukraine" = ukrain_tags,
                    "Hamas" = hamas_tags)

for(event in names(events_tags)) {
  for(a_tag in levels(all_tags$Tag)) {
    comparison_for_binomial(events_tags[[event]], event, a_tag)
  }  
}
```

Cumulative binomial
```{r}
binomuial_cum_model <- cmdstan_model("stan/binomial_cum.stan")


comparison_for_cumulative_binomuial <- function(tags_df, event, tag){
  crnt_tags_df <-
    tags_df |>
    filter(Tag == tag)
  
  max_days <- 
    crnt_tags_df |>
    select(Period, AbsDay) |>
    group_by(AbsDay) |>
    count() |>
    filter(n == 2) |>
    pull(AbsDay) |>
    max()
  
  
  cum_df <-
    crnt_tags_df |>
    filter(AbsDay > 0, AbsDay <= max_days) |>
    group_by(Period) |>
    arrange(AbsDay) |>
    mutate(CumN = cumsum(N),
           CumNtotal = cumsum(Ntotal)) |>
    ungroup() |>
    mutate(AbsDay = factor(AbsDay, levels = 1:max_days)) |>
    complete(Period, AbsDay) |>
    group_by(Period) |>
    fill(CumN, CumNtotal, .direction = "down") |>
    ungroup() |>
    mutate(AbsDay = as.integer(AbsDay)) |>
    arrange(Period, AbsDay) |>
    mutate(P = CumN / CumNtotal,
           LowerCI = binom::binom.exact(CumN, CumNtotal)$lower,
           UpperCI = binom::binom.exact(CumN, CumNtotal)$upper)
  

    
  binom_cum_stan_data <- list(
    DaysN = max_days,
    Ntag = cum_df$CumN,
    Ntotal = cum_df$CumNtotal,
    Period= as.integer(cum_df$Period),
    AbsDay = cum_df$AbsDay)
  
  binom_cum_fit <- binomuial_cum_model$sample(binom_cum_stan_data, chains = 4, parallel_chains = future::availableCores())
  binom_cum_draws <- binom_cum_fit$draws(format = "df") |> as_tibble()
  
  binom_cum_posterior <-
    binom_cum_draws |>
    select(.draw, starts_with("p")) |>
    pivot_longer(starts_with("p"), names_to = "Term", values_to = "P") |>
    mutate(Indexes = str_extract_all(Term, "\\d+"),
           AbsDay = purrr::map_int(Indexes, ~as.integer(.[[1]])),
           Period = factor(purrr::map_chr(Indexes, ~.[[2]]), labels = levels(crnt_tags_df$Period))) |>
    select(-Term, -Indexes)
  
  binom_cum_posterior_avg <-
    binom_cum_posterior|>
    group_by(Period, AbsDay) |>
    summarise(Avg = median(P),
              LowerCI = kilter::lower_ci(P),
              UpperCI = kilter::upper_ci(P),
              .groups = "drop")
  
  binom_cum_diff <-
    binom_cum_posterior |>
    pivot_wider(names_from = Period, values_from = P) |>
    group_by(AbsDay) |>
    summarise(P = mean(Post - Pre > 0), .groups = "drop")
  
  ggplot(cum_df, aes(x = AbsDay, y = P, color = Period)) +
    geom_hline(yintercept = c(0.9, 0.95), linetype = "dashed") +
    geom_ribbon(data = binom_cum_posterior_avg, aes(y = Avg, ymin = LowerCI, ymax = UpperCI, fill = Period), color = NA, alpha = 0.35) +
    geom_point(data = binom_cum_diff, color = "black") +
    geom_errorbar(aes(ymin = LowerCI, ymax = UpperCI)) +
    geom_line() +
    geom_point() +
    scale_color_manual(values = colors_pre_post) +
    scale_fill_manual(values = colors_pre_post) +
    scale_y_continuous(name = "Probability of message", sec.axis = dup_axis(name = NULL, breaks = c(0.9, 0.95))) +
    scale_x_continuous(name = "Day") +
    labs(title = glue::glue("{event} - {tag}"))
  
  ggsave(glue("figures/cumulative-{event}-{tag}.png"), units = "cm", width = 24, height = 14)
}

for(event in names(events_tags)) {
  for(a_tag in levels(all_tags$Tag)) {
    comparison_for_cumulative_binomuial(events_tags[[event]], event, a_tag)
  }  
}
```


### Multinomial
```{r}
categorical_model <- cmdstan_model("stan/categorical.stan")

  crnt_tags_df <-
    tags_df |>
    filter(!(Tag %in% c("War", "Ecology"))) |>
    mutate(Tag = fct_drop(Tag))

  max_days_in_period <- c("Pre" = abs(min(crnt_tags_df$Day)), "Post" =  max(crnt_tags_df$Day))
  max_days <- min(max_days_in_period)
  
  cat_stan_data <- list(
    N = nrow(crnt_tags_df),
    TagsN = nlevels(crnt_tags_df$Tag),
    Tag = as.integer(crnt_tags_df$Tag),
    Day = as.integer(crnt_tags_df$Day)
  )
  
  cat_fit <- categorical_model$sample(cat_stan_data, chains = 4, parallel_chains = future::availableCores())
  cat_draws <- cat_fit$draws(format = "df") |> as_tibble()
  saveRDS(cat_draws, "draws-categorical-ukraine.RDS")
```

```{r}
  # extracting alpha and beta for linear models
  alpha2 <-
    kilter::parse_stan_posterior_array(cat_draws, "alpha", list("Tag" = levels(crnt_tags_df$Tag)[-1], "Period" = c("Pre", "Post")), "alpha") |>
    mutate(Tag = as.character(Tag))
  alpha <- bind_rows(
    alpha2, 
    alpha2 |>
      group_by(.chain, .draw, Period) |>
      summarise(.groups = "drop") |>
      mutate(alpha = 0, Tag = levels(crnt_tags_df$Tag)[1])) |>
  mutate(Tag = factor(Tag, levels = levels(crnt_tags_df$Tag)),
         Period = factor(Period, levels = levels(crnt_tags_df$Period)))

  beta2 <- 
    kilter::parse_stan_posterior_array(cat_draws, "beta", list("Tag" = levels(crnt_tags_df$Tag)[-1], "Period" = c("Pre", "Post")), "beta") |>
    mutate(Tag = as.character(Tag))
  beta <- bind_rows(
    beta2, 
    beta2 |>
      group_by(.chain, .draw, Period) |>
      summarise(.groups = "drop") |>
      mutate(beta = 0, Tag = levels(crnt_tags_df$Tag)[1])) |>
  mutate(Tag = factor(Tag, levels = levels(crnt_tags_df$Tag)),
         Period = factor(Period, levels = levels(crnt_tags_df$Period)))
  ab <- 
    left_join(alpha, beta, by = c(".chain", ".draw", "Tag", "Period")) |>
    select(-.chain) |>
    arrange(.draw, Period, Tag)

  compute_categorical_p_single_day <- function(params_df, day){
    params_df |>
      mutate(score = exp(alpha + beta * day),
             P = score / sum(score)) |>
      select(.draw, Tag, P) |>
      mutate(AbsDay = day)
  }
  
  compute_categorical_p <- function(params_df, period) {
    purrr::map(0:max_days_in_period[period], ~compute_categorical_p_single_day(params_df, .)) |>
      list_rbind()
  }

  ab_pre <- filter(ab, Period == "Pre")
  categorical_predictions_pre <- purrr::map(1:max(ab_pre$.draw), ~filter(ab_pre, .draw == .) |> compute_categorical_p("Pre"), .progress = TRUE) |> list_rbind()
  categorical_predictions_pre$Period <- "Pre"
  saveRDS(categorical_predictions_pre, "cat-predictions-pre-ukraine.RDS")

  ab_post <- filter(ab, Period == "Post")
  categorical_predictions_post <- purrr::map(1:max(ab_post$.draw), ~filter(ab_post, .draw == .) |> compute_categorical_p("Post"), .progress = TRUE) |> list_rbind()
  categorical_predictions_post$Period <- "Post"
  saveRDS(categorical_predictions_post, "cat-predictions-post-ukraine.RDS")

  categorial_predictions <- 
    bind_rows(categorical_predictions_pre, categorical_predictions_post) |>
    mutate(Period = factor(Period, levels = c("Pre", "Post")))
  saveRDS(categorial_predictions, "cat-predictions-ukraine.RDS")
  
  categorial_predictions_avg <-
    categorial_predictions |>
    group_by(Tag, Period, AbsDay) |>
    summarise(Avg = mean(P),
              LowerCI = kilter::lower_ci(P),
              UpperCI = kilter::upper_ci(P),
              .groups = "drop")
  
  categorial_diff <-
    categorial_predictions |>
      group_by(Tag, AbsDay, Period) |>
      mutate(Iteration = 1:n()) |>
      pivot_wider(values_from = P, names_from = Period) |>
      summarise(P = mean(Post > Pre), .groups = "drop")
    
  
    
  ggplot(crnt_tags_df, aes(x = AbsDay, y = P, color = Period)) +
    geom_hline(yintercept = c(0.9, 0.95), linetype = "dashed") +
    geom_point(data = categorial_diff, color = "black") +
    geom_ribbon(data = categorial_predictions_avg, aes(y = Avg, ymin = LowerCI, ymax = UpperCI, fill = Period), color = NA, alpha = 0.35) +
    geom_line(data = categorial_predictions_avg, aes(y = Avg)) +
    geom_point() +
    facet_wrap(. ~ Tag) +
    scale_x_continuous(name = "Day") +
    scale_y_continuous(name = "Probability of message", sec.axis = dup_axis(name = NULL, breaks = c(0.9, 0.95))) +
    scale_color_manual(values = colors_pre_post) +
    scale_fill_manual(values = colors_pre_post) +
    labs(title = glue::glue("{event}"))


    
```



Fixed period
```{r}
period_days_n <- 14

ukrain_tags |>
  mutate(Delta = Datum - ukrain_start) |>
  filter(Datum >= ukrain_start - as.difftime(period_days_n, units = "days"),
         Datum <= ukrain_start + as.difftime(period_days_n, units = "days")) |>
ggplot(aes(x = Datum, y = P, color = PreWar)) + 
  geom_point(show.legend = FALSE) +
  facet_wrap(. ~ Tag, scales = "free_y")  

```



```{r}
message_complete_df_long |>
  group_by(Datum, pre_ukr, pre_hamas) |>
  summarise(TotalWar = sum(War),
            N = n(),
            .groups = "drop") |>
ggplot(aes(x = Datum, y = N, color = pre_hamas)) +
  geom_point()
```
